### 2020-12-08 Thesis-Chapter1-Introduction.tex
Spelling and editorial improvements
Address some TODOs
Address more todos

Certain policy issues are \acp{wicked-prob}. Originally coming from the field of design theory, the term
``\ac{wicked-prob}'' uses ``wicked'' not in the moral sense, but in the malignant, vicious, and tricky sense
\cite{rittel_dilemmas_1973}. Unlike their ``tame'' counterparts which science, engineering, and traditional policymaking
are well equipped to answer, \acp{wicked-prob} lack clear formulations, causes, resolutions, or measurements. Each
attempted solution has permanent and often unintended consequences, and likely exists in a pattern of chronic policy
failure.

@@

% TODO: change EA everywhere to "lawful access"?

@@

\principle{\Ac{EA} is an inherently complex problem}{The factors factors that make \ac{EA} a \ac{wicked-prob} mean that
proposed solutions must be analyzed at many levels. Some these factors are \ac{EA}'s relation to mass surveillance,
potential for abuse, international consequences, economic impact, and need for transparency.}

>>>>

Certain policy issues are \acp{wicked-prob}. Originally coming from the field of design theory, the term
``\ac{wicked-prob}'' uses ``wicked'' not in the moral sense, but in the malignant, vicious, and tricky sense
\cite{rittel_dilemmas_1973}. Unlike their ``tame'' counterparts which science, engineering, and traditional policymaking
are well equipped to answer, \acp{wicked-prob} lack clear formulations, causes, resolutions, or measurements. Each
attempted solution has permanent and often unintended consequences, and is likely to exist in a pattern of chronic
policy failure.

@@

@@

\principle{\Ac{EA} is an inherently complex problem}{The factors factors that make \ac{EA} a \ac{wicked-prob} mean that
proposed solutions must be analyzed at many levels. Some these factors are \ac{EA}'s relation to \ac{masssurv},
potential for abuse, international consequences, economic impact, and need for transparency.}

### 2020-12-08 Thesis-Chapter3-PolicyMaking.tex
Address more todos

\Acp{wicked-prob} always consist of subproblems, which may be tame or wicked in and of themselves. For example, climate
change has the tame subproblem of developing clean energy technology, and the wicked subproblem of implementing any
solution in a way that doesn't leave vast swaths of people behind.

% TODO: what am I really trying to express here?
% ... WPs quickly grow in complexity with the number of subproblems that comprise them. Refining the problem statement
% and combining tame solutions with wicked subproblem mitigations is the truly ``wicked'' part.

% Elements of the problem may be tame---for example, one can independently confirm that reducing global carbon emissions
% by a certain amount will result in a certain difference in response with a certain statistical confidence. Other
% elements of the problem are wicked---for example, one cannot verify the effectiveness of a plan ahead of time, port a
% solution from a different domain, or definitively evaluate its performance after the fact.

% Refining the problem statement and combining the
% tame elements into a cohesive solution is the ``wicked'' part.

@@

Rozenshtein reflects on the diagnosis optimistically:

@@

% FIXME: include this idea somewhere in this section.
% Note: a problem with muddling through wicked problems is that each attempted fix leads to long lasting unintended side
% effects. This is especially true with matters of encryption, where the data could be stored for years till it can be
% decrypted! So must be very careful to avoid solution attempts that leave long scars.

@@

Let us define one strategy for handling \ac{wicked-prob} as corralling it back into the classes of problems we are
better at dealing with. That means breaking it into tame sub-problems and developing agreeable strategies for its
irreducibly wicked elements. Once that is done, we can rationally analyze and increment our way to victory. There is no
handbook for how to do this, but several sources offer advice.

@@

% TODO: Dulimarta: consider changing this heading to "Proposal: Observe-Orient-Decide-Act"
\section{Proposal: The OODA Loop for Wicked Problems}

@@

The \ac{OODAloop} was a developed by Air Force Colonel John Boyd as a description of a successful strategy for
countering opponents in real-time combat \cite{angerman_2004}. Boyd created it in the context of military strategy, but
its ideas have penetrated many other sectors, including cybersecurity, where it serves as a model for structuring
incident response \cite{schneier_future_2014}. There is an emphasis on fast cycle times and ``getting inside'' your
opponent's loop as a means of overcoming raw power with speed and agility. It is illustrated in \myfig{fig-ooda-loop}.

@@

% TODO: Kalafut: It's not clear to me that this paragraph is a good fit here. It is good content. Perhaps this paragraph
% or its ideas should be at the start of the Argument Maps section.

@@

Traditional policymaking typically follows either a rational approach rooted in the scientific method, aiming to be
thorough and complete, or an intuitional approach rooted in evolutionary trial and error, humbly aiming only to take
steps in the right direction. Both approaches have their strengths, but neither is suited to the complex and
controversial nature of so-called ``\acp{wicked-prob}.'' The strategy proposed here relies on the goodwill participation
of all parties, who engage in ongoing research and debate that rejects easy answers, encourages flexible risk-based
solutions, and crystallizes discussion around specific proposals. \Acp{argmap} are one tool that may facilitate this
process. The result from a single iteration is an imperfect step forward based on implementing a particular proposal.
The result from many iterations is the breakdown of the problem into sub-problems---some tame and inevitably some still
wicked. The tame sub-problems can be addressed with the \ac{classical-method}, and the wicked sub-problems, by now
restrained under a sound and agreeable strategy, can be addressed with \ac{incrementalism}. This process is demonstrated
in \myfig{fig-policy-ooda-process}.

% TODO: explain the icons

>>>>

% Elements of the problem may be tame---for example, one can develop a new renewable energy widget and another can
% independently confirm its efficiency. Other elements of the problem are wicked---for example, one cannot predict the
% side effects of the deploying the widget, and there is now way to know how effort spent there could have been used
% elsewhere. \Acp{wicked-prob} always consist of subproblems, which may be tame or wicked in and of themselves

\Acp{wicked-prob} always consist of subproblems, which may be tame or wicked in and of themselves. For example, climate
change has the tame subproblem of developing clean energy technology, and the wicked subproblem of implementing any
solution in a way that doesn't leave vast swaths of people behind. \Acp{wicked-prob} quickly grow in complexity with the
number of subproblems that comprise them. Truly imposing \acp{wicked-prob} are a tangle of contributing factors, and
despite the presence of tame elements, they pose a large number of difficult challenges.

% ... WPs quickly grow in complexity with the number of subproblems that comprise them. Refining the problem statement
% and combining tame solutions with wicked subproblem mitigations is the truly ``wicked'' part.

% Elements of the problem may be tame---for example, one can independently confirm that reducing global carbon emissions
% by a certain amount will result in a certain difference in response with a certain statistical confidence. Other
% elements of the problem are wicked---for example, one cannot verify the effectiveness of a plan ahead of time, port a
% solution from a different domain, or definitively evaluate its performance after the fact. Refining the problem
% statement and combining the tame elements into a cohesive solution is the ``wicked'' part.

% Elements of the problem may be tame---one can independently confirm that reducing global carbon emissions by a certain
% amount will result in a certain difference in response with a certain statistical confidence. \Acp{wicked-prob} always
% consist of subproblems, which may be tame or wicked in and of themselves. Refining the problem statement and combining
% the tame elements in a cohesive solution is the ``wicked'' part. Despite certain tame elements, one cannot verify the
% effectiveness of a plan ahead of time, port a solution from a different domain, or even definitively evaluate its
% performance after the fact.

@@

\Ac{encryption} technology is particularly sensitive to the property that policy implementation has irreversible
effects. Technology deployments have long tails, and attackers have the ability to record and store data for later
analysis. In one investigation, Australian authorities cracked a cold case based on evidence acquired from a mobile
device that they cracked five years after it was seized \cite{evans_arrests_2020}. Lawful authorities wouldn't be the
only ones to benefit from a policy miscalculation could negatively impact security for a very long time.

Rozenshtein reflects on the \ac{wicked-prob} diagnosis optimistically:

@@

@@

Let us define one strategy for handling \ac{wicked-prob} as corralling it back into the classes of problems we are
better at dealing with. That means breaking it into tame subproblems and developing agreeable strategies for its
irreducibly wicked elements. Once that is done, we can rationally analyze and increment our way to victory. There is no
handbook for how to do this, but several sources offer advice.

@@

\section{Proposal: The OODA Loop for Wicked Problems}

@@

The \acl{OODAloop} was a developed by Air Force Colonel John Boyd as a description of a successful strategy for
countering opponents in real-time combat \cite{angerman_2004}. Boyd created it in the context of military strategy, but
its ideas have penetrated many other sectors, including cybersecurity, where it serves as a model for structuring
incident response \cite{schneier_future_2014}. There is an emphasis on fast cycle times and ``getting inside'' your
opponent's loop as a means of overcoming raw power with speed and agility. It is illustrated in \myfig{fig-ooda-loop}.

@@

% TODO: Kalafut: It's not clear to me that this paragraph is a good fit here. It is good content. Perhaps this paragraph
% or its ideas should be at the start of the Argument Maps section.
% PLAN: I will move this to when I talk more about argument maps. After the refactor, they won't even be introduced yet
% here.

@@

Traditional policymaking typically follows either a rational approach rooted in the scientific method, aiming to be
thorough and complete, or an intuitional approach rooted in evolutionary trial and error, humbly aiming only to take
steps in the right direction. Both approaches have their strengths, but neither is suited to the complex and
controversial nature of so-called ``\acp{wicked-prob}.'' The strategy proposed here relies on the goodwill participation
of all parties, who engage in ongoing research and debate that rejects easy answers, encourages flexible risk-based
solutions, and crystallizes discussion around specific proposals. \Acp{argmap} are one tool that may facilitate this
process. The result from a single iteration is an imperfect step forward based on implementing a particular proposal.
The result from many iterations is the breakdown of the problem into subproblems---some tame and inevitably some still
wicked. The tame subproblems can be addressed with the \ac{classical-method}, and the wicked subproblems, by now
restrained under a sound and agreeable strategy, can be addressed with \ac{incrementalism}.

% TODO: do I need this explanation?

This process is demonstrated in \myfig{fig-policy-ooda-process}. An out of control problem (the large storm) is being
confronted. Several iterations of the \ac{OODAloop} gradually bring it down to size, spinning out tame problems (puzzle
pieces) addressed by classical analysis and wicked but restrained problems (small storms) controlled by
\ac{incrementalism}.

### 2020-12-08 Thesis-Chapter4-Arguments.tex
Clean up glossary
Updates based on presentation feedback
Address more todos

\myfig{fig-args-contrib} maps the central arguments in the EA debate. As previously stated, it is \ac{encryption}'s dual
role as enabler of radical privacy and cornerstone of security is at the heart of the debate. Of course, there is only a
conflict if one sees radical privacy as a bad thing. The right to privacy is both a strongly held value and an enshrined
legal principle, and many use privacy-based arguments to show \ac{EA} as socially undesirable. The \ac{snowden}
revelations \cite{landau_making_2013} only unveiled the scale of privacy-eroding U.S. government surveillance enabled by
technological changes, terrorism-motivated policies, and weak oversight \cite{shamsi_2011}. Violations of privacy are
concerning for those most vulnerable and for society at large. For example, Saudi journalist Jamal Khashoggi's private
communications were surveilled by the Saudi government before he became a victim of politically motivated murder
\cite{liebermann_2019}. At a societal level, presence of surveillance alone changes behavior and chills free speech
\cite{rogaway_moral_2015}, while government violations of the law degrade institutional trust.

@@

% FIXME: I need to address \cite{koepke_2020} in these maps.

@@

The argument has a few blind spots. Depicting privacy and safety as competing in a zero-sum tradeoff ignores all the
ways in which privacy directly enhances safety. Data privacy makes certain crimes more difficult, such as stalking or
identity theft. Privacy is also a public good to the extent that it precludes mass surveillance, an argument explored
more deeply in the ``Golden Age'' side of the argument.

Most importantly, the ``Going Dark'' argument hinges on encryption being responsible for failure of law enforcement
investigations. As described in \mysec{sed-arg-contrib}, there is a lack of data on the extent to which encryption is
itself to blame for failed cases. One can certainly understand the argument, but the lack of evidence is an issue. The
highest-profile case of \ac{the-cw2}, Apple vs. FBI, ended with the \ac{FBI} breaking into the phone but finding nothing
of value. Tellingly, each of Comey's examples in his 2014 speech were not cases where encryption inhibited
investigations, but where cases hypothetically \ii{could} have been inhibited if encryption was a factor
\cite{comey_2014}.

Citing lack of evidence only gets you so far, however. In most cases, it is impossible to know what would have happened
in an investigation unimpeded by strong encryption. This is part of why it is a \ac{wicked-prob}. Indeed, when the
encryption curtain is pulled back, there is often much to see. Unencrypted Facebook Messenger is has approximately 20\%
market share in active monthly users \cite{statista_2020} but accounts for 65\% of \ac{CSAM} reports
\cite{keller_internet_2019}, and European law enforcement's infiltration of encrypted messaging service Encrochat
exposed crime rings and led to a massive raids \cite{cox_2020}.

Aside from encryption entirely, a 2018 study found that law enforcement's primary challenges with digital evidence were
identifying service providers and getting the sought after data from them \cite{carter_2018}. This is in the context of
current legal frameworks and access capabilities. The report is fittingly titled ``Low-Hanging Fruit,'' and also lists
resource limitations and training needs among top priorities. Indeed, many current initiatives are underfunded. Major
anti-\ac{CSAM} legislation passed in 2008, but has been funded at only 50\% of authorized levels, the Justice Department
task force has produced only two of five biennial reports, and a senior executive position within the Justice Department
was never created \cite{keller_internet_2019}. Federal entities commissioned to improve communication between law
enforcement and industry as well as to train law enforcement on managing digital evidence are similarly underfunded
\cite{carter_2018}.

@@

The argument for law enforcement having \ii{too much} access to data claims that instead of law enforcement ``Going
Dark,'' they are living in the ``Golden Age for Surveillance'' \cite{swire_encryption_2011}. This argument goes that
although encryption makes certain data completely unaccessible, technological change on the whole has given authorities
much more digital evidence than encryption has taken away. Paired with the dangers of mass surveillance, encryption is
a necessary defense, and any \ac{EA} implementation would compromise encryption's protection. The argument is mapped in
\myfig{fig-arg-golden-age}.

@@

The ``Golden Age'' argument revolves around surveillance, its undesirability, and technology's role in it. Cryptographer
Phillip Rogaway writes at length about the negative impacts of mass surveillance in his essay ``The Moral Character of
Cryptographic Work'' \cite{rogaway_moral_2015}. Rogaway argues that ``pervasive collection \ii{itself} chills
free-speech and threatens liberal democracy'' (emphasis his) through the way its mere existence affects the subject. He
argues that it is fundamentally a tool of power, and though I would add that power from the people created through
representative political process is valid power, clear abuses have undercut reasonable confidence in these institutions.

The next element is that technology has increased the government's surveillance powers. This can hardly be contested.
Going back to Comey's ``Going Dark'' examples once more, each case he cited was cracked by digital evidence that would
not have been available even twenty years ago \cite{comey_2014}. In response to the danger of mass surveillance and the
scale of modern surveillance capabilities, encryption is one of the only privacy-saving tools. (Opting out may be
possible for some, but is not a realistic option for society as a whole.)

While much stronger than the ``Going Dark'' argument, the ``Golden Age for Surveillance'' argument suffers from one
major flaw. It focuses strongly on the abilities of and dangers posed by agencies like the \ac{NSA} and ignores the
problems of local, state, and sometimes-surprisingly-cash-strapped federal law enforcement. The boom in the existence of
digital evidence is clear, but as shown, law enforcement already has trouble acquiring the evidence with the help of
tech and communications companies, and does not have the funding or capability to get it on its own \cite{carter_2018}.
Exacerbating the situation is the ``tech effect,'' which is increasing jurors' expectations of hard evidence in cases
where it may be suspected to exist \cite{shelton_study_2006}.

Still, the fact that the tools of mass surveillance are out of reach for most law enforcement must not be relied upon.
That is, most likely, a temporary situation. A safer criticism of the ``Golden Age'' framing of \ac{EA} is that \ac{EA}
without mass surveillance is conceivable. \ac{EA} implementations that enable law enforcement but thwart mass
surveillance by remote powers subvert most of the ``Golden Age'' argument. Means for \ac{EA} to do so are explored in
future argument maps.

@@

@@

The final two anti-\ac{EA} fallacies are examples of false dichotomies, presenting the situation as if there are only
two choices. \Acp{wicked-prob} have an unlimited set of possible responses, so reducing it to two (and often, presenting
one as a straw man) is misleading. One false dichotomy is that the only acceptable solutions are perfect \ac{EA} or no
\ac{EA}. This leaves no room for risk-based approaches, the typical security strategy outside the realm of cryptography.
Another false dichotomy is that regulators must either allow strong cryptography or we will live in a mass surveillance
dystopia. This one carries hints of ``slippery slope'' and appeal to emotion as well, framed as a choice between only
two futures. Framing the situation in all or nothing terms harms the chances for successful collaborative debate.

@@

The availability and cost of exploits ebbs and flows. On one hand, January 2019 sported a \$2 million price on the
exploit market for an Apple iOS jailbreak bug \cite{goodin_zeroday_2019}. On the other hand, an October 2020 report
shows that law enforcement agencies can extract data from nearly every mobile phone on the market for less than \$2000 a
unit \cite{koepke_2020}. As the exploit arms race continues, exploitability will be inconsistent. In the end, law
enforcement would face periods where it struggles for access, and would need to develop costly exploits in the lab.
Mandatory vulnerability disclosure is key to a \ac{lawful-hacking} proposal; however, based on the \ac{FBI}'s historical
abuses of power \cite{shamsi_2011} and Congress's distaste for funding even initiatives it has passed
\cite{keller_internet_2019} or providing strong oversight \cite{johnson_congressional_2004}, one cannot imagine that the
government would actually give up expensive vulnerabilities easily when they are the pillar of their access strategy.

The final proposal is \ac{EA} itself. Its advantage, generally speaking, is that it would provide reliable access to law
enforcement in a reliable manner. It could serve as a compliment to \ac{lawful-hacking}, blunting the incentive to hoard
vulnerabilities because it would be unnecessary for \ac{EA} compliant systems, which would presumably include mainstream
technology. In that situation, \ac{EA} would combat common crime and \ac{lawful-hacking} would combat organized crime
and terror groups. While impact on civil liberties and mass surveillance depends on implementation, any proposal would
be costly to introduce and threaten U.S. product competitiveness due to perceived insecurity. Indeed, risk to security
is the largest weakness of \ac{EA}. Any proposal has to overcome significant challenges to provide acceptable levels of
risk. The next map analyzes this more closely.

@@

% FIXME: add note about how device security actually totally sucks

Finally, one \ac{DAR}-specific \ac{EA} class is device key escrow. In this scheme, key recovery material exists on the
device itself. While not necessary to classify as device key escrow, most of these proposals include physical possession
requirements and additionally ``brick'' the device or make it absolutely clear that \ac{EA} has been performed
\cite{savage_lawful_2018} \cite{ozzie_2018}. Such features preclude mass surveillance and ensure transparency to the
device user. This approach relies on secure hardware and a method of authorizing the recovery request to the device
itself. The authentication step can be paired with other \ac{EA} strategies, such as engaging with trusted-parties or
distributed systems. This is where the device escrow shows its weaknesses. Secure hardware enclaves, though improving,
are not totally reliable, and authentication to the hardware device is susceptible to the same attacks as traditional
escrow \ac{EA}.

>>>>

\myfig{fig-args-contrib} maps the central arguments in the EA debate. As previously stated, it is \ac{encryption}'s dual
role as enabler of radical privacy and cornerstone of security is at the heart of the debate. Of course, there is only a
conflict if one sees radical privacy as a bad thing. The right to privacy is both a strongly held value and an enshrined
legal principle, and many use privacy-based arguments to show \ac{EA} as socially undesirable. The Snowden revelations
\cite{landau_making_2013} only unveiled the scale of privacy-eroding U.S. government surveillance enabled by
technological changes, terrorism-motivated policies, and weak oversight \cite{shamsi_2011}. Violations of privacy are
concerning for those most vulnerable and for society at large. For example, Saudi journalist Jamal Khashoggi's private
communications were surveilled by the Saudi government before he became a victim of politically motivated murder
\cite{liebermann_2019}. At a societal level, presence of surveillance alone changes behavior and chills free speech
\cite{rogaway_moral_2015}, while government violations of the law degrade institutional trust.

@@

@@

The argument has a few blind spots. Depicting privacy and safety as competing in a zero-sum tradeoff ignores all the
ways in which privacy directly enhances safety. Data privacy makes certain crimes more difficult, such as stalking or
identity theft. Privacy is also a public good to the extent that it precludes \ac{masssurv}, an argument explored more
deeply in the ``Golden Age'' side of the argument.

Most importantly, the ``Going Dark'' argument hinges on encryption being responsible for failure of law enforcement
investigations. As described in \mysec{sed-arg-contrib}, there is a lack of data on the extent to which encryption is
itself to blame for failed cases. One can certainly understand the argument, but the lack of evidence is an issue. The
highest-profile case of \ac{the-cw2}, Apple vs. FBI, ended with the \ac{FBI} breaking into the phone but finding nothing
of value. Tellingly, each of Comey's examples in his 2014 speech were not cases where encryption inhibited
investigations, but where cases hypothetically \ii{could} have been inhibited if encryption was a factor
\cite{comey_2014}. More recently, an 2020 report shows that law enforcement agencies currently have the upper hand on
mobile device encryption, and are actually able to extract data from nearly every mobile phone on the market
\cite{koepke_2020}.

Citing lack of evidence that encryption is stymieing cases only gets you so far, however. In most cases, it is
impossible to know what would have happened in an investigation unimpeded by strong encryption. This is part of why it
is a \ac{wicked-prob}. Indeed, when the encryption curtain is pulled back, there is often much to see. Unencrypted
Facebook Messenger is has approximately 20\% market share in active monthly users \cite{statista_2020} but accounts for
65\% of \ac{CSAM} reports \cite{keller_internet_2019}, and European law enforcement's infiltration of encrypted
messaging service Encrochat exposed crime rings and led to a massive raids \cite{cox_2020}.

Aside from the challenge of encryption entirely, a 2018 study found that law enforcement's primary difficulties with
digital evidence were identifying service providers and getting the sought after data from them \cite{carter_2018}. This
is in the context of current legal frameworks and access capabilities. The report is fittingly titled ``Low-Hanging
Fruit,'' and also lists resource limitations and training needs among top priorities. Indeed, many current initiatives
are underfunded. Major anti-\ac{CSAM} legislation passed in 2008, but has been funded at only 50\% of authorized levels,
the Justice Department task force has produced only two of five biennial reports, and a senior executive position within
the Justice Department was never created \cite{keller_internet_2019}. Federal entities commissioned to improve
communication between law enforcement and industry as well as to train law enforcement on managing digital evidence are
similarly underfunded \cite{carter_2018}.

@@

The argument for law enforcement having \ii{too much} access to data claims that instead of law enforcement ``Going
Dark,'' they are living in the ``Golden Age for Surveillance'' \cite{swire_encryption_2011}. This argument goes that
although encryption makes certain data completely unaccessible, technological change on the whole has given authorities
much more digital evidence than encryption has taken away. Paired with the dangers of \ac{masssurv}, encryption is a
necessary defense, and any \ac{EA} implementation would compromise encryption's protection. The argument is mapped in
\myfig{fig-arg-golden-age}.

@@

The ``Golden Age'' argument revolves around surveillance, its undesirability, and technology's role in it. Cryptographer
Phillip Rogaway writes at length about the negative impacts of \ac{masssurv} in his essay ``The Moral Character of
Cryptographic Work'' \cite{rogaway_moral_2015}. Rogaway argues that ``pervasive collection \ii{itself} chills
free-speech and threatens liberal democracy'' (emphasis his) through the way its mere existence affects the subject. He
argues that it is fundamentally a tool of power, and though I would add that power from the people created through
representative political process is valid power, clear abuses have undercut reasonable confidence in these institutions.

The next element is that technology has increased the government's surveillance powers. This can hardly be contested.
Going back to Comey's ``Going Dark'' examples once more, each case he cited was cracked by digital evidence that would
not have been available even twenty years ago \cite{comey_2014}. In response to the danger of \ac{masssurv} and the
scale of modern surveillance capabilities, encryption is one of the only privacy-saving tools. (Opting out may be
possible for some, but is not a realistic option for society as a whole.)

While much stronger than the ``Going Dark'' argument, the ``Golden Age for Surveillance'' argument suffers from one
major flaw. It focuses strongly on the abilities of and dangers posed by agencies like the \ac{NSA} and ignores the
problems of local, state, and sometimes-surprisingly-cash-strapped federal law enforcement. The boom in the existence of
digital evidence is clear, but as shown, law enforcement already has trouble acquiring many forms of evidence with the
help of tech and communications companies, and does not have the funding or capability to get it on its own
\cite{carter_2018}. Exacerbating the situation is the ``tech effect,'' which is increasing jurors' expectations of hard
evidence in cases where it may be suspected to exist \cite{shelton_study_2006}.

Still, the fact that the tools of \ac{masssurv} are out of reach for most law enforcement must not be relied upon. That
is, most likely, a temporary situation. Law enforcement's present ability to access mobile devices proves this point. A
safer criticism of the ``Golden Age'' framing of \ac{EA} is that \ac{EA} without \ac{masssurv} is conceivable. \ac{EA}
implementations that enable law enforcement but thwart \ac{masssurv} by remote powers subvert most of the ``Golden Age''
argument. \ac{EA} proposals striving for these properties are explored in future argument maps.

@@

% TODO: Dulimarta: Orange/green boxes for consistency? Decide when I make a key.

@@

The final two anti-\ac{EA} fallacies are examples of false dichotomies, presenting the situation as if there are only
two choices. \Acp{wicked-prob} have an unlimited set of possible responses, so reducing it to two (and often, presenting
one as a straw man) is misleading. One false dichotomy is that the only acceptable solutions are perfect \ac{EA} or no
\ac{EA}. This leaves no room for risk-based approaches, the typical security strategy outside the realm of cryptography.
Another false dichotomy is that regulators must either allow strong cryptography or we will live in a \ac{masssurv}
dystopia. This one carries hints of ``slippery slope'' and appeal to emotion as well, framed as a choice between only
two futures. Framing the situation in all or nothing terms harms the chances for successful collaborative debate.

@@

The availability and cost of exploits ebbs and flows. On one hand, January 2019 sported a \$2 million price on the
exploit market for an Apple iOS jailbreak bug \cite{goodin_zeroday_2019}. On the other hand, the October 2020 report
detailing law enforcement's mobile device cracking capabilities estimated the per-device cost to be only \$2000 a unit
\cite{koepke_2020}. As the exploit arms race continues, exploitability will be inconsistent. In the end, law enforcement
would face periods where it struggles for access, and would need to develop costly exploits in the lab. Mandatory
vulnerability disclosure is key to a \ac{lawful-hacking} proposal; however, based on the \ac{FBI}'s historical abuses of
power \cite{shamsi_2011} and Congress's distaste for funding even initiatives it has passed \cite{keller_internet_2019}
or providing strong oversight \cite{johnson_congressional_2004}, one cannot imagine that the government would actually
give up expensive vulnerabilities easily when they are the pillar of their access strategy.

The final proposal is \ac{EA} itself. Its advantage, generally speaking, is that it would provide reliable access to law
enforcement in a reliable manner. It could serve as a compliment to \ac{lawful-hacking}, blunting the incentive to hoard
vulnerabilities because it would be unnecessary for \ac{EA} compliant systems, which would presumably include mainstream
technology. In that situation, \ac{EA} would combat common crime and \ac{lawful-hacking} would combat organized crime
and terror groups. While impact on civil liberties and \ac{masssurv} depends on implementation, any proposal would be
costly to introduce and threaten U.S. product competitiveness due to perceived insecurity. Indeed, risk to security is
the largest weakness of \ac{EA}. Any proposal has to overcome significant challenges to provide acceptable levels of
risk. The next map analyzes this more closely.

@@

Finally, one \ac{DAR}-specific \ac{EA} class is device key escrow. In this scheme, key recovery material exists on the
device itself. While not necessary to classify as device key escrow, most of these proposals include physical possession
requirements and additionally ``brick'' the device or make it absolutely clear that \ac{EA} has been performed
\cite{savage_lawful_2018} \cite{ozzie_2018}. Such features preclude \ac{masssurv} and ensure transparency to the device
user. This approach relies on secure hardware and a method of authorizing the recovery request to the device itself. The
authentication step can be paired with other \ac{EA} strategies, such as engaging with trusted-parties or distributed
systems. This is where the device escrow shows its weaknesses. Secure hardware enclaves, though improving, are not
totally reliable, and authentication to the hardware device is susceptible to the same attacks as traditional escrow
\ac{EA}.

### 2020-12-08 Thesis-Chapter6-PathsAhead.tex
Address more todos

However, at the same time as encryption is enabling strong privacy, the broader technological path society is taking is
stripping it away. The ubiquity and insecurity of networked computers has been a bear to privacy, enabling threats
ranging from identity theft, to inappropriate data mining, to mass surveillance. To the extent that encryption mitigates
these, it increases public safety.

@@

In this thesis I propose a method to contend with \acp{wicked-prob}, a modified \ac{OODAloop} that emphasizes
collaborative debate between individuals diverse in background and values, facilitated in part by argument maps, and
focusing on specific proposals. I try to capture that here by surveying the debate, mapping the arguments, and
ultimately performing threat model analysis of a specific \ac{EA} proposal for \acl{DAR}. I now turn to conclusions and
paths forward for both the technical and policymaking communities.

@@

% TODO: make a \term for mass surveillance

The additions to the threat model and the instability of present circumstances mean that additional research into
\ac{EA} is fully justified. It belongs as just a component of the effort that society should be putting towards
addressing the broader set of issues. Fortunately, developments in secure hardware mark real progress in the search for
acceptably secure \ac{EA}. Secure hardware technology is clearly imperfect, and alone does not resolve the numerous
challenges that full-fledged \ac{EA} proposals must overcome, but device escrow for \acl{DAR} such as that described in
Savage's proposal offers transparency, protection from mass surveillance, and \acl{LE} utility while making progress on
security.

% FIXME: "The conclusions I draw do not settle the issue..." Make sure wording is good.

>>>>

However, at the same time as encryption is enabling strong privacy, the broader technological path society is taking is
stripping it away. The ubiquity and insecurity of networked computers has been a bear to privacy, enabling threats
ranging from identity theft, to inappropriate data mining, to \ac{masssurv}. To the extent that encryption mitigates
these, it increases public safety.

@@

In this thesis I propose a method to confront \acp{wicked-prob}, a modified \ac{OODAloop} that emphasizes collaborative
debate between individuals diverse in background and values, facilitated in part by argument maps, and focusing on
specific proposals. I try to capture that here by surveying the debate, mapping the arguments, and ultimately performing
threat model analysis of a specific \ac{EA} proposal for \acl{DAR}. I now turn to conclusions and paths forward for both
the technical and policymaking communities.

@@

The additions to the threat model and the instability of present circumstances mean that additional research into
\ac{EA} is fully justified. It belongs as just a component of the effort that society should be putting towards
addressing the broader set of issues. Fortunately, developments in secure hardware mark real progress in the search for
acceptably secure \ac{EA}. Secure hardware technology is clearly imperfect, and alone does not resolve the numerous
challenges that full-fledged \ac{EA} proposals must overcome, but device escrow for \acl{DAR} such as that described in
Savage's proposal offers transparency, protection from \ac{masssurv}, and \acl{LE} utility while making progress on
security.
